### 2017-02-28 grlp.tex
further changes

%!TEX root =  autocontgrlp.tex
\begin{comment}
\section{Approximate Linear Programming: Successes and Challenges}
When the MDP has a large number of states, it is difficult to solve for $J^*$ using either the linear program \eqref{mdplp} or other full state representation methods such as value iteration or policy iteration \cite{BertB}. A practical solution is to resort to function approximation. Linear function approximation, wherein the solution is searched in the subspace spanned by the column vectors of a given feature matrix $\Phi$.
The approximate linear program (ALP) is obtained by making use of LFA in the LP, i.e., by introducing the new variables $r\in \R^k$ and adding the extra constraint $J=\Phi r$ in \eqref{mdplp} with $\Phi \in \R^{n\times k}$ \citep{SchSei85}.
By substitution, this leads to
\begin{align}\label{alp}
\begin{split}
\min_{r\in \R^k}\, &c^\top \Phi r\,\,\,
\text{s.t.}\mb \Phi r\geq T \Phi r,
\end{split}
\end{align}
where $J\geq TJ$ is a shorthand for the $nd$ constraints in \eqref{mdplp} and $\Phi$ is a feature matrix whose first column is $\one$. Unless specified otherwise we use $\tr$ to denote an arbitrary solution to the ALP, and we let $\tj=\Phi \tr$ to denote the corresponding approximate value function and $\tu$ to denote the greedy policy w.r.t. $\tj$.
The following is a preliminary error bound for the ALP from \cite{ALP}:
\begin{theorem}[Error Bound for ALP]
\begin{align*}
||J^*-\tj||_{1,c}\leq \frac{2}{1-\alpha}\min_{r}||J^*-\Phi r||_\infty
\end{align*}

@@

where $P_{u^*}$ is the probability transition matrix under $u^*$ ($P_{u^*}(s,s') = P_{u^*(s)}(s,s')$, $s,s'\in S$) \cite{dolgov}. Even though we might not have the optimal policy $u^*$ in practice, the fact that $\lambda^*$ is a probability distribution and that it is a linear combination of $\{P_{u^*},P^2_{u^*},\ldots\}$ hints at the kind of features that might be useful for the $W$ matrix.

@@

\input{cartoon}
\Cref{cartoon} which shows the solutions to the LP, ALP and GRLP respectively. The error in the ALP solution has already been studied in \cite{ALP}. Our objective is to study the extra source of error due to constraint approximation. We now define two projection operators which are central to our error analysis and in them we assume that the set $\N'\subset \R^k$ is such that $\N' = \N + t \one$ for any $t\in \R$.

@@

Then, for $J\in \R^n$, $\Gamma J$,
the \emph{ least upper bound projection} of $J$ is defined as

@@

Then, the \emph{approximate least upper bound (ALUB) projection operator}

@@

$\F_J$ contains vectors in the span of $\Phi$ that upper bound $TJ$. Further, since $(\Gamma J)(i) = \min\{ V(i) \,:\, V\in \F_J \}$, it also follows that $ V\ge \Gamma J $ holds for any $V\in \F_J$.

@@

$||J^*-\hj||_{1,c}
\leq\frac{1}{1-\alpha}(6~\us{\min}{r\in \R^k}||J^*-\Phi r||_{\infty}
+2||\Gamma J^*-\hg J^*||_{\infty})$.

@@

\leq 2\left(\frac{1}{(1-\alpha)^2}\right)\, \big( 2~\us{\min}{r\in \R^k} ||J^*-\Phi r||_{\infty}
+||\Gamma J^*-\hg J^*||_{\infty}+||\hj-\hg\hj||_{\infty}\big)$.

@@

\textbf{On Prediction Error:}The first factor in the right hand side of the prediction error in \Cref{cmt2} is related to the best possible approximation that can be achieved with the chosen feature matrix $\Phi$. This term is inherent to the ALP formulation as shown in \Cref{alpvanilla}. The second factor in the right hand side of the prediction error is related to constraint approximation and is completely defined in terms of $\Phi$, $W$ and $T$, and does not require knowledge of stationary distribution of the optimal policy. It makes intuitive sense since given that $\Phi$ approximates $J^*$, it is enough for $W$ to depend on $\Phi$ and $T$ without any additional requirements.\par

@@

&\leq 2\left(\frac{c^\top \psi}{(1-\beta_{\psi})^2}\right)\, \big( 2||J^*-\Phi r^*||_{\infty}

@@

+||\Gamma J^*-\hg J^*||_{\infty}+||\hj-\hg\hj||_{\infty}\big).

@@

\begin{theorem}[Constraint Sampling]\label{st}
Let $s\in S$ be a state whose constraint is selected by $W$ (i.e., for some $i$ and all $(s',a)\in S\times A$,

@@

It is important to note that \Cref{rlpt} holds only in high probability and is valid only under idealized assumption of knowing the optimal policy $u^*$ and \Cref{cmt2} does not have these limitations.
In addition, from \Cref{st}, we know that the additional error $|\Gamma J^*(s) -\hg J^*(s)|$ due to constraint approximation is less than the original projection error $|\Gamma J^*(s)-J^*(s)|$ due to function approximation. This means that for the RLP to perform well it is enough to retain those states for which the linear function approximation via $\Phi$ is known to perform well.

>>>>

%!TEX root =  autocontgrlp.tex
\section{Approximate Linear Programming: Successes and Challenges}
When the MDP has a large number of states, it is difficult to solve for $J^*$ using either the linear program \eqref{mdplp} or other full state representation methods such as value iteration or policy iteration \cite{BertB}. A practical solution is to resort to function approximation. Linear function approximation, wherein the solution is searched in the subspace spanned by the column vectors of a given feature matrix $\Phi$.
The approximate linear program (ALP) is obtained by making use of LFA in the LP, i.e., by introducing the new variables $r\in \R^k$ and adding the extra constraint $J=\Phi r$ in \eqref{mdplp} with $\Phi \in \R^{n\times k}$ \citep{SchSei85}.
By substitution, this leads to
\begin{align}\label{alp}
\min_{r\in \R^k}\, &c^\top \Phi r\,\,\,
\text{s.t.}\mb \Phi r\geq T \Phi r,
where $J\geq TJ$ is a shorthand for the $nd$ constraints in \eqref{mdplp} and $\Phi$ is a feature matrix whose first column is $\one$. Unless specified otherwise we use $\tr$ to denote an arbitrary solution to the ALP, and we let $\tj=\Phi \tr$ to denote the corresponding approximate value function and $\tu$ to denote the greedy policy w.r.t. $\tj$.
The following is a preliminary error bound for the ALP from \cite{ALP}:
\begin{theorem}[Error Bound for ALP]
\norm{J^*-\tj}_{1,c}\leq \frac{2}{1-\alpha}\min_{r}\norm{J^*-\Phi r}_\infty

@@

where $P_{u^*}$ is the probability transition matrix under $u^*$ ($P_{u^*}(s,s') = P_{u^*(s)}(s,s')$, $s,s'\in S$) \cite{dolgov}. Even though we might not have the optimal policy $u^*$ in practice, the fact that $\lambda^*$ is a probability distribution and that it is a linear combination of $\{P_{u^*},P^2_{u^*},\ldots\}$ hints at the kind of features that might be useful for constructing the $W$ matrix.

@@

%\input{cartoon}
\begin{figure}
\includegraphics[scale=0.7]{cartoon_grlp.pdf}
\end{figure}
\Cref{cartoon} which shows the solutions to the LP, ALP and GRLP respectively. The error in the ALP solution has already been studied in \cite{ALP}. Our objective is to study the extra source of error due to constraint approximation. We now define two projection operators that are central to our error analysis and in them we assume that the set $\N'\subset \R^k$ is such that $\N' = \N + t \one$ for any $t\in \R$.

@@

For $J\in \R^n$, $\Gamma J$, the \emph{least upper} (LU) projection of $J$ is defined as

@@

The \emph{approximate least upper} (ALU) projection operator

@@

$\F_J$ contains all vectors in the span of $\Phi$ that upper bound $TJ$. Further, since $(\Gamma J)(i) = \min\{ V(i) \,:\, V\in \F_J \}$, it also follows that $ V\ge \Gamma J $ holds for any $V\in \F_J$. Also $\Gamma J\geq T J$.

@@

$\norm{J^*-\hj}_{1,c}
\leq\frac{1}{1-\alpha}(6~\us{\min}{r\in \R^k}\norm{J^*-\Phi r}_{\infty}
+2\norm{\Gamma J^*-\hg J^*}_{\infty})$.

@@

\leq 2\left(\frac{1}{(1-\alpha)^2}\right)\, \big( 2~\us{\min}{r\in \R^k} \norm{J^*-\Phi r}_{\infty}
+\norm{\Gamma J^*-\hg J^*}_{\infty}+\norm{\hj-\hg\hj}_{\infty}\big)$.

@@

\textbf{On Prediction Error:} The first factor in the right hand side of the prediction error in \Cref{cmt2} is related to the best possible approximation that can be achieved with the chosen feature matrix $\Phi$. This term is an carry over of the upper bound in ALP formulation as shown in \Cref{alpvanilla}. The second factor in the right hand side of the prediction error is related to constraint approximation and is completely defined in terms of $\Phi$, $W$ and $T$, and does not require knowledge of stationary distribution of the optimal policy.\par

@@

&\leq 2\left(\frac{c^\top \psi}{(1-\beta_{\psi})^2}\right)\, \big( 2\norm{J^*-\Phi r^*}_{\infty}

@@

+norm{\Gamma J^*-\hg J^*}_{\infty}+\norm{\hj-\hg\hj}_{\infty}\big).

@@

\begin{corollary}[Constraint Sampling]\label{st}
$W\in \{0,1\}^{nd\times m}$Let $s\in S$ be a state whose constraint is selected by $W$ (i.e., for some $i$ and all $(s',a)\in S\times A$,

@@

\end{corollary}
It is important to note that \Cref{rlpt} holds only in high probability and is valid only under idealized assumption of knowing the optimal policy $u^*$, while \Cref{cmt2} does not have these limitations.
In addition, the error $|\Gamma J^*(s) -\hg J^*(s)|$ (in \Cref{st} ) due to constraint approximation is less than the original projection error $|\Gamma J^*(s)-J^*(s)|$ due to function approximation. This means that for the RLP to perform well it is important to retain the constraints corresponding to those states for which the linear function approximation via $\Phi$ is known to perform well.

### 2017-02-28 intro.tex
further changes

The so-called dynamic programming (DP) compute the optimal value function $J^*$ first and then obtain an optimal policy $u^*$ using $J^*$. Conventional DP  techniques, such as value-, or policy-iteration, or linear programming (LP) \cite{BertB} can compute the exact value of $J^*$ (and $u^*$). However, a shortcoming of these conventional methods is that their computational overhead grows with the number of states, a practical hindrance when the MDP has a large number of states.\par
In this paper we consider the problem of using dynamic programming with MDPs with large state spaces. A practical way to tackle the issue of large number of states is to compute an approximate value function $\tilde{J}$ instead of $J^*$, and a sub-optimal policy $\tu$ can be obtained using $\tj$\footnote{It is computationally cheap to obtain such a $\tu$ from $tj$ or $u^*$ from $J^*$}. The sub-optimal policy can then be used to make the decision and results in a cumulative return given by $J_{\tu}$. The idea here is forsake optimality to gain computational tractability. The success then depends on the quality of approximation, i.e., on the quantities $\norm{J^*-\tilde{J}}$ and $\norm{J^*-J_{\tu}}$ ($\norm{\cdot}$ is an appropriate norm) known as prediction error and the control error respectively.\par
Linear function approximation (LFA), i.e., letting $\tilde{J}=\Phi r^*$ where $\Phi$ is a so-called feature matrix and $r^*$ is a weight vector to be computed, is the most widely used method of approximation. Here, dimensionality reduction is achieved by choosing $\Phi$ to have fewer columns in comparison to the number of states, holding the promise of being able to work with MDPs regardless of the number of states.\par
The \emph{approximate linear program} (ALP) \cite{ALP,CS,SALP,ALP-Bor} and its variants introduce linear function approximation in the linear programming formulation. A critical shortcoming of vanilla ALP is that the number of constraints are of the order of the size of the state space, making this vanilla version intractable. A way out is to choose a subset of constraints at random and drop the rest, thereby formulating a \emph{reduced linear program} (RLP). The performance analysis of the RLP can be found in \cite{CS} and the RLP has also been shown to perform well in experiments \cite{ALP,CS,CST}. An alternative approach to handle the issue of large number of constraints is to employ function approximation in the dual variables of the ALP \cite{ALP-Bor,dolgov}, an approach that was also found useful in experiments. However, to this date, there exist no theoretical guarantees bounding the loss in performance resulting from such an approximation.

@@

The \emph{approximate linear program} (ALP) \cite{ALP,CS,SALP,ALP-Bor} and its variants introduce LFA in the linear programming formulation to dynamic programming. Theoretical performance analysis of the ALP can be found in \cite{ALP}.
%and a salient feature is that it does not suffer from issues such as `policy oscillations'\footnote{ADP methods that use linear least squares projection are known to exhibit `policy oscillations' %\cite{BertB}, i.e., output a repeating sequence of bad sub-optimal policies. Since our focus in this paper is the ALP formulation, we refrain from a detailed presentation of the other methods.}
The number of variables of the ALP is only $k$, a critical shortcoming of the ALP is that the number of constraints are of the order of the size of the state space, making it intractable.
Two approaches have been found empirically successful \cite{CS,dolgov,ALP-Bor} in addressing the issue of large number of constraints in the ALP. In the first approach \cite{CS}, a random subset of constraints is chosen (dropping the rest), thereby formulating a \emph{reduced linear programming} (RLP) problem. The performance analysis of the RLP can be found in \cite{CS}, however, the bounds hold only in high probability under idealized assumptions. The second approach involves employing function approximation in the dual variables of the ALP \cite{ALP-Bor,dolgov}. However, to this date, there exist no theoretical guarantees bounding the loss in performance due to this approach.\par

Our motivation stems from the fact that ALP with tractable number of constraints will result in a full dimeniosnality free ADP method. However, constraint reduction in the ALP is an extra source of error (in addition to the error due to LFA), which has not been theoretically well understood.  The focus of this paper is to fill this gap in theory by deriving performance bounds for constraint reduction in the ALP formulation.
The salient aspects of our contribution are listed below:
\begin{enumerate}
\item We define a generalized reduced linear programming (GRLP) which has a tractable number of constraints that are obtained as positive linear combinations of the original constraints. The GRLP amounts to linear function approximation of the dual variables, and the RLP is a special case of GRLP.
\item We develop a novel analytical machinery to bound the prediction error $\norm{J^*-\hj}$ where $hj$ is the solution to the GRLP.
		\item We also bound the performance loss due to using the policy $\hu$ that is one-step greedy with respect to $\hj$ (Theorem~\ref{polthe}).
		\item Our analysis is based on two novel $\max$-norm contraction operators and our results hold \emph{deterministically}, as opposed to the results on RLP \cite{SALP,CS}, where the guarantees have a probabilistic nature.
%Our analysis also makes use of arguments based on \emph{Lyapunov} functions, an approach much similar to prior works in the ALP literature \cite{ALP,SALP}.
\item Our results on GRLP are the first to theoretically analyze the use of linear function approximation of Lagrangian (dual) variables underlying the constraints.
\item A numerical example in controlled queues is provided to illustrate the theory.
\end{enumerate}
%ALP with tractable number of constraints would result in  a full dimensionality free ADP method, without issues such as policy oscillations.
A short and preliminary version of this paper without the theoretical analysis is available in \cite{aaaipaper}. The rest of the paper is organized as follows:
\end{comment}

>>>>

The so-called dynamic programming (DP) methods \cite{BertB} compute the optimal value function $J^*$ first and then obtain an optimal policy $u^*$ using $J^*$\footnote{Obtaining $u^*$ from $J^*$ is computationally cheap}. Conventional DP  techniques, such as value-, or policy-iteration, or linear programming (LP) \cite{BertB} can compute the exact value of $J^*$ (and $u^*$). However, a shortcoming of these conventional methods is that their computational overhead grows with the number of states, a practical hindrance when the MDP has a large number of states.\par
In this paper we consider the problem of using dynamic programming with MDPs with large state spaces. A practical way to tackle the issue of large number of states is to compute an approximate value function $\tilde{J}$ instead of $J^*$, and a sub-optimal policy $\tu$ can be obtained using $\tj$\footnote{It is computationally cheap to obtain such a $\tu$ from $\tj$ or $u^*$ from $J^*$}. The sub-optimal policy can then be used to make the decision and results in a cumulative return given by $J_{\tu}$. The idea here is forsake optimality to gain computational tractability. The success then depends on the quality of approximation, i.e., on the quantities $\norm{J^*-\tilde{J}}$ and $\norm{J^*-J_{\tu}}$ ($\norm{\cdot}$ is an appropriate norm) known as prediction error and the control error respectively.\par
Linear function approximation (LFA), i.e., letting $\tilde{J}=\Phi r^*$ where $\Phi$ is a so-called feature matrix and $r^*$ is a weight vector to be computed, is the most widely used method of approximation. Here, dimensionality reduction is achieved by choosing $\Phi$ to have fewer columns in comparison to the number of states, holding the promise of being able to work with MDPs regardless of the number of states. Note that the matrix $\Phi$ is implicit in these methods: the full matrix is neither stored nor computed, but only its rows are accessed as needed.\par
The \emph{approximate linear program} (ALP) \cite{ALP,CS,SALP,ALP-Bor} and its variants introduce linear function approximation in the linear programming formulation. A critical shortcoming of vanilla ALP is that the number of constraints are of the order of the size of the state space, making this vanilla version intractable in MDPs with large number of states. A way out is to choose a subset of constraints at random and drop the rest, thereby formulating a \emph{reduced linear program} (RLP). The performance analysis of the RLP can be found in \cite{CS} and the RLP has also been shown to perform well in experiments \cite{ALP,CS,CST}. An alternative approach to handle the issue of large number of constraints is to employ function approximation in the dual variables of the ALP \cite{ALP-Bor,dolgov}, an approach that was also found useful in experiments. However, to this date, there exist no theoretical guarantees for this approach.\par

@@

The \emph{approximate linear program} (ALP) \cite{ALP,CS,SALP,ALP-Bor} and its variants introduce LFA in the linear programming formulation to dynamic programming. Theoretical performance analysis of the ALP can be found in \cite{ALP}.
%and a salient feature is that it does not suffer from issues such as `policy oscillations'\footnote{ADP methods that use linear least squares projection are known to exhibit `policy oscillations' %\cite{BertB}, i.e., output a repeating sequence of bad sub-optimal policies. Since our focus in this paper is the ALP formulation, we refrain from a detailed presentation of the other methods.}
The number of variables of the ALP is only $k$, a critical shortcoming of the ALP is that the number of constraints are of the order of the size of the state space, making it intractable.
Two approaches have been found empirically successful \cite{CS,dolgov,ALP-Bor} in addressing the issue of large number of constraints in the ALP. In the first approach \cite{CS}, a random subset of constraints is chosen (dropping the rest), thereby formulating a \emph{reduced linear programming} (RLP) problem. The performance analysis of the RLP can be found in \cite{CS}, however, the bounds hold only in high probability under idealized assumptions. The second approach involves employing function approximation in the dual variables of the ALP \cite{ALP-Bor,dolgov}. However, to this date, there exist no theoretical guarantees bounding the loss in performance due to this approach.\par
Our motivation stems from the fact that ALP with tractable number of constraints will result in a full dimeniosnality free ADP method. However, constraint reduction in the ALP is an extra source of error (in addition to the error due to LFA), which has not been theoretically well understood.  The focus of this paper is to fill this gap in theory by deriving performance bounds for constraint reduction in the ALP formulation.
\item We define a generalized reduced linear programming (GRLP) which has a tractable number of constraints that are obtained as positive linear combinations of the original constraints. The GRLP amounts to linear function approximation of the dual variables, and the RLP is a special case of GRLP.
\item We develop a novel analytical machinery to bound the prediction error $\norm{J^*-\hj}$ where $hj$ is the solution to the GRLP.
		\item We also bound the performance loss due to using the policy $\hu$ that is one-step greedy with respect to $\hj$ (Theorem~\ref{polthe}).
%Our analysis also makes use of arguments based on \emph{Lyapunov} functions, an approach much similar to prior works in the ALP literature \cite{ALP,SALP}.
\item Our results on GRLP are the first to theoretically analyze the use of linear function approximation of Lagrangian (dual) variables underlying the constraints.
%ALP with tractable number of constraints would result in  a full dimensionality free ADP method, without issues such as policy oscillations.
A short and preliminary version of this paper without the theoretical analysis is available in \cite{aaaipaper}. The rest of the paper is organized as follows:
\end{comment}

### 2017-02-28 mdp.tex
addressed the comments by Csaba
further changes

\label{bellpol}u^*(s)&= \underset{a\in A}{\argmax}\big(g_a(s)+\alpha \us{\sum}{s'\in S}p_a(s,s')J^*(s')\big),\,.

@@

We now present some definition which will be useful later:

@@

\item The Bellman operator $T\colon \R^n \ra \R^n$ is given by $(TJ)(s)=\max_{a \in A}\big(g_a(s)+\alpha \sum_{s'} p_a(s,s')J(s')\big).

@@

\begin{align*} u_J(s)\eqdef\underset{a\in A}{\argmax}\big(g_a(s)+\alpha \us{\sum}{s'\in S} p_a(s,s')J(s')\big).\end{align*}

@@

||J||_{1,c}=\sum_{s \in S} c(s)|J(s)|$.

@@

and $\norm{\cdot}_{\infty,\rho}$, $\parallel J\parallel_{\infty}=\max_{s\in S}|J(s)|$ and $||J||_{\infty,\rho}=\max_{s \in S} \frac{|J(s)|}{\rho(s)}$.

>>>>

\label{bellpol}u^*(s)&\in \underset{a\in A}{\argmax}\big(g_a(s)+\alpha \us{\sum}{s'\in S}p_a(s,s')J^*(s')\big),\,.

@@

For the subsequent sections, the following definitions will be useful later:

@@

\item\label{bellopval} The Bellman operator $T\colon \R^n \ra \R^n$ is given by $(TJ)(s)=\max_{a \in A}\big(g_a(s)+\alpha \sum_{s'} p_a(s,s')J(s')\big).

@@

\begin{align*} u_J(s)\in\underset{a\in A}{\argmax}\big(g_a(s)+\alpha \us{\sum}{s'\in S} p_a(s,s')J(s')\big).\end{align*}

@@

\norm{J}_{1,c}=\sum_{s \in S} c(s)|J(s)|$.

@@

and $\norm{\cdot}_{\infty,\rho}$, $\norm{J}_{\infty}=\max_{s\in S}|J(s)|$ and $\norm{J}_{\infty,\rho}=\max_{s \in S} \frac{|J(s)|}{\rho(s)}$.

### 2017-02-28 num.tex
further changes

We set $c(s)=(1-\zeta) \zeta^s$, where $s=1,\ldots,9999$, with $\zeta=0.9$ and $\zeta=0.999$ respectively. The results in Table~\ref{pref} show that the performance exhibited by $W_a$ and $W_c$ is better by several orders of magnitude over `random' in the case of the large system $Q_L$ and is close to that of the ideal sampler $W_i$. Note that computing $\et$ was hard in the case of large $n=10^4$ and since $\et$ is completely dependent on the structure of $\Phi$, $T$ and $W$ we instead computed the same for $n=10$ and used it as a surrogate. Accordingly, we chose a smaller system $Q_S$ with $n=10$, $d=2$, $k=2$, $m=5$, $q(1)=0.2$, $q(2)=0.4$, $p=0.2$ and $\alpha=0.98$. Note that a better performance of $W_a$ and $W_c$ is in agreement with a lower value of $\et$.

>>>>

We set $c(s)=(1-\zeta) \zeta^s$, where $s=1,\ldots,9999$, with $\zeta=0.9$ and $\zeta=0.999$ respectively. The results in Table~\ref{pref} show that the performance exhibited by $W_a$ and $W_c$ is better by several orders of magnitude over `random' in the case of the large system $Q_L$ and is close to that of the ideal sampler $W_i$. Note that computing $\et$ was expensive in the case of large $n=10^4$ and since $\et$ is completely dependent on the structure of $\Phi$, $T$ and $W$ we instead computed the same for $n=10$ and used it as a surrogate. Accordingly, we chose a smaller system $Q_S$ with $n=10$, $d=2$, $k=2$, $m=5$, $q(1)=0.2$, $q(2)=0.4$, $p=0.2$ and $\alpha=0.98$. Note that a better performance of $W_a$ and $W_c$ is in agreement with a lower value of $\et$.

### 2017-02-28 prob.tex
further changes

\section{Linear Programming for MDP}
In this section, we review relevant results \cite{BertB,ALP,CS} in the linear programming approach to MDPs. Throughout the rest of the paper, we let $c\in (0,1)^n$ to denote a probability distribution with strictly positive components.

@@

\textbf{Approximate Linear Programming (ALP)} is obtained by making use of LFA in the LP, i.e., by introducing the new variables $r\in \R^k$ and adding the extra constraint $J=\Phi r$ in \eqref{mdplp} with $\Phi \in \R^{n\times k}$ \citep{SchSei85}. By substitution, this leads to

@@

where $E\Phi r\geq H \Phi r$ is a shorthand for the $nd$ constraints in \eqref{mdplp} which uses the definition of $H$ from \Cref{notations}-\eqref{bellactval}. Alternatively, we also use the notation $\Phi r \geq T\Phi r$, instead of $E\Phi r\geq H \Phi r$ using the definition of $T$ from \Cref{notations}-\eqref{bellopval}. Note that, for $k=n$ and $\Phi=I_{n\times n}$, the ALP is same as the LP. In what follows, we use the notation $\tj=\Phi \tr$ and $\tu=u_{\tj}$ where $\tr\in\R^k$ denotes an arbitrary solution to ALP in \eqref{alp}.\par

@@

Let $\one\in \R^k$ belong to the column span of $\Phi$ and let $\mu_u\eqdef (1-\alpha)c^\top\top(I-\alpha P_u)^{-1}$ be a probability distribution over $S$ induced by a policy $u$. Then it follows that

@@

\item\label{alppe} The prediction error is bounded as $\norm{J^*-\tj}_{1,c}\leq \frac{2}{1-\alpha}\min_{r}\norm{J^*-\Phi r}_\infty$.
\item\label{alpce} The control error is bounded as $\norm{J^*-J_{\tu}}_{1,c}\leq \frac{2}{1-\alpha}\norm{J^*-\tj}_{1,\mu_{\tu}}$.

@@

The gist of \Cref{alpvanilla}-\eqref{alppe} is that the prediction error $\norm{J^*-\tj}_{1,c}$ of ALP is worse utmost by a constant factor than the best possible error that can be achieved using the span of $\Phi$. Further from \Cref{alpvanilla}-\eqref{alpce} the control error is worse utmost by a constant factor than the prediction error.\par

@@

where $\I\subset S \times A$ is the index set of sampled constraints, i.e., $|\I|=m$ and $\N$\footnote{The appendix explains how $\N$ can be chosen analytically.} is a compact set such that $\tr \in \N$, and it ensures boundedness of the RLP \eqref{alp}. We denote an arbitrary solution to the RLP by $\tr_{RLP}$.\par
We now state a useful result from \cite{CS} that characterizes the performance analysis RLP as follows:

@@

The gist of \Cref{rlpt} is that when `sufficient' number of samples are retained using a specialized distribution ($\psi_{u^*}$), then with high probability, the prediction error of RLP and ALP differ only by a small factor (i.e., $\epsilon\norm{J^*}_{1,c}$).\par
A major weakness of \Cref{rlpt} is that it holds only for RLP formulated using $\psi_{u^*}$ which is impractical since $u^*$ is unknown in general. However, empirical studies \cite{ALP,CS,CST} have shown that RLP formulated with other sampling distributions do perform well in experiments. In this context, another interesting approach is linear approximation of the dual variables \cite{dolgov}, which also is known to work in experiments but without any theoretical guarantees. These fact hint to the fact that the theoretical understanding of RLP and constraint sampling/approximation needs to be improved. Further, such theoretical understanding will justify that ALP with constraint reduction is a fully dimensionality free method in that the problem variables as well as constraints are independent of the number of states of the MDP.

>>>>

\section{Exact and Approximate Linear Programming for MDP}
In this section, we review relevant results \cite{BertB,ALP,CS} in the linear programming approach to MDPs. Throughout the rest of the paper, we let $c\in (0,1)^n$ to denote a probability distribution with strictly positive components.\\

@@

Once $J^*$ is obtained from \eqref{mdplp}, an optimal policy $u^*$ can then be obtained via \eqref{bellpol} for each state relatively cheaply even for large state-spaces. However, notice that \eqref{mdplp} is a linear program in $n$ variables with $nd$ constraints. As a result, when $n$ is large it is computationally expensive to solve \eqref{mdplp}, a fact that motivated the need to look at the approximate linear programming (ALP) formulation which computes an approximate value function $\tj$.\\
\textbf{Approximate Linear Programming (ALP)} is obtained by adding the extra constraint $J=\Phi r$ in \eqref{mdplp} with $\Phi \in \R^{n\times k}$ \citep{SchSei85}. By substitution, this leads to

@@

where $E\Phi r\geq H \Phi r$ is a shorthand for the $nd$ constraints in \eqref{mdplp} which uses the definition of $H$ from \Cref{notations}-\eqref{bellactval}.
ALP introduces what is known as linear function approximation in the LP, i.e., the value function computed by ALP can be given by $\tj=\Phi \tr$ for an appropriate $\tr\in \R^k$. ALP is a linear program in $k$ ($<<n$) variables and hence is computationally easier to handle in practice than LP. Note that, for $k=n$ and $\Phi=I_{n\times n}$, the ALP is same as the LP. In what follows, we use the notation $\tj=\Phi \tr$ and $\tu=u_{\tj}$ (the greedy policy with respect to $\tj$, see \Cref{notations}-\eqref{greedy}) where $\tr\in\R^k$ denotes an arbitrary solution to ALP in \eqref{alp}. Further, when convenient, we also use the notation $\Phi r \geq T\Phi r$, instead of $E\Phi r\geq H \Phi r$ in \eqref{alp} using the definition of $T$ from \Cref{notations}-\eqref{bellopval}.\\
\textbf{Error Analysis of ALP:} ALP is an approximate method as opposed to LP. Thus, it is important for us to characterize the approximation error. In the context of ALP for MDPs, we care about two error terms namely the \emph{prediction error} $\norm{\tj-J^*}$, that captures the error in the approximate value function $\tj$ and the \emph{control error} $\norm{J_{\tu}-J^*}$, that is the loss in performance when a sub-optimal policy $\tu$ is used instead of an optimal policy $u^*$ (note that here $\norm{\cdot}$ is an appropriate norm).\par

@@

Let $\one\in \R^k$ belong to the column span of $\Phi$ and let $\mu_u\eqdef (1-\alpha)c^\top(I-\alpha P_u)^{-1}$ be a probability distribution over $S$ induced by a policy $u$. Then it follows that

@@

\item\label{alppe} $\norm{J^*-\tj}_{1,c}\leq \frac{2}{1-\alpha}\min_{r}\norm{J^*-\Phi r}_\infty$.
\item\label{alpce} $\norm{J^*-J_{\tu}}_{1,c}\leq \frac{2}{1-\alpha}\norm{J^*-\tj}_{1,\mu_{\tu}}$.

@@

The gist of \Cref{alpvanilla}-\eqref{alppe} is that the prediction error $\norm{J^*-\tj}_{1,c}$ of ALP is worse utmost by a constant factor than the best possible error that can be achieved using the span of $\Phi$. Further from \Cref{alpvanilla}-\eqref{alpce} the control error is worse utmost by a constant factor than the prediction error.\\
\textbf{Issues with ALP}: While ALP has only $k$ variables, it has $nd$ constraints.
Two approaches have been found empirically successful \cite{CS,dolgov,ALP-Bor} in addressing the issue of large number of constraints in the ALP. In the first approach \cite{CS}, a random subset of constraints is chosen (dropping the rest), thereby formulating a \emph{reduced linear programming} (RLP) problem. The performance analysis of the RLP can be found in \cite{CS}, however, the bounds hold only in high probability under idealized assumptions. The second approach involves employing function approximation in the dual variables of the ALP \cite{ALP-Bor,dolgov}. However, to this date, there exist no theoretical guarantees bounding the loss in performance due to this approach. We now dicuss RLP and its error analysis.\\

@@

where $\I\subset S \times A$ is the index set of sampled constraints, i.e., $|\I|=m$ and $\N$\footnote{The appendix explains how $\N$ can be chosen analytically.} is a compact set such that $\tr \in \N$, and it ensures boundedness of the RLP \eqref{alp}. We denote an arbitrary solution to the RLP by $\tr_{RLP}$.\\
\textbf{Error Analysis of RLP:} We now state a useful result from \cite{CS} that characterizes the performance analysis RLP as follows:

@@

The gist of \Cref{rlpt} is that when `sufficient' number of samples are retained using a specialized distribution ($\psi_{u^*}$), then with high probability, the prediction error of RLP and ALP differ only by a small factor (i.e., $\epsilon\norm{J^*}_{1,c}$).\\
\textbf{Limitations of Error Analysis for RLP:} A major weakness of \Cref{rlpt} is that it holds only for RLP formulated using $\psi_{u^*}$ which is impractical since $u^*$ is unknown in general. However, empirical studies \cite{ALP,CS,CST} have shown that RLP formulated with other sampling distributions do perform well in experiments. In this context, another interesting approach is linear approximation of the dual variables \cite{dolgov}, which also is known to work in experiments but without any theoretical guarantees. These fact hint to the fact that the theoretical understanding of RLP and constraint sampling/approximation needs to be improved. Further, such theoretical understanding will justify that ALP with constraint reduction is a fully dimensionality free method in that the problem variables as well as constraints are independent of the number of states of the MDP.
